Deep learning for quality control of surface physiographic fields using satellite Earth observations

### 2023-05-29 docs/submission/response_to_reviewer_comments/reviewer_1/response_reviewer_1.tex

\documentclass[10pt]{article}
\usepackage{vmargin}
\setpapersize{USletter}
\usepackage{ragged2e}

\author{Tom Kimpson \\ 
	\and Margarita Choulga\\ 
 \and Matthew Chantry\\
\and Gianpaolo Balsamo\\  
 \and Souhail Boussetta\\
  \and Peter Dueben\\
   \and Tim Palmer\\
}
\title{\normalsize Response to Reviewer 1 Comments
  concerning HESS submission egusphere-2022-1177}
\begin{document}
	\maketitle
\noindent Author response to referee reports for the paper egusphere-2022-1177,
entitled \textit{``Deep Learning for Verification of Earth-System Parametrisation of Water Bodies"}. \newline

\noindent We thank the referee for their reading, helpful criticism and suggestions towards the improvement of the manuscript. \newline

\noindent We have addressed each point in turn below and the manuscript has been updated accordingly \newline

\noindent  We hope that this satisfies the request for changes necessary to proceed with the publication of the updated manuscript. \newline

\section*{Essential comments}

\begin{enumerate}
	\item \textbf{Essential comment 1: mixing of different models, parameters and predictors}

	We agree with the general points here raised by the reviewer that the terminology used was insufficiently precise. This has now been corrected throughout the entire manuscript. We now make it clear that we are always comparing the results between two neural network models, not comparing NN predictions with e.g. FLake. If we update some input field to our NN and the NN prediction accuracy improves, this is good evidence that the updated field itself is more accurate and would enable a physical model like FLake to make more accurate predictions.

	Taking some individual points:

	\begin{itemize}
		\item \textbf{L.79, L82-87}. What was meant here is the distinction between is the new information informative enough to have visible changes, or it is just lots of work for no impact? This has now been properly phrased in the text. Whilst we did briefly discuss local degradation previously we have also now included a more explicit discussion. Additionally, we explicitly calculate the training noise by retraining VESPER multiple times. We now highlight the effect of the training noise in the updated manuscript. Having retrained the model multiple times, our conclusions are generally unchanged for all grid-point categories, with the exception of the Vegetation category which has significant training noise over a small number of grid cells making it difficult to draw meaningful conclusions. This is again discussed in the text and we thank the reviewer for raising this point for our attention. 		
		\item \textbf{2.2 L88-89} As the point above, we have retrained VESPER multiple times to show that the training noise is generally smaller than the prediction changes due to the different input fields. We reword the text throughout to be clear that we are always comparing two NN models.
		\item \textbf{2.3 L105-106}  As above, terminology and text has been changed throughout the manuscript to be more precise
		\item \textbf{2.4 L19}  As discussed we have now trained multiple versions of the model to better quantify the training noise and our conclusions remain unchanged that we can use VESPER to (a) check that an updated field is closer to reality and (b) see if this updated field increases the accuracy of our NN model. Both of these points are relevant for the updated fields within a physical model like FLake.
	\end{itemize}

		\item \textbf{Essential comment 2: strange results and speculative explanations}

		As mentioned we now explicitly the explicitly calculate the training noise by retraining VESPER multiple times. For the points in Northern Canada, the Toshka lakes and the Vegetation category our previously quoted changes are less than the training noise. Again we thank the reviewer for highlighting this to us. We have removed the discussion on  Northern Canada and the Toshka lakes from the manuscript, whilst the effect of the noise on the Vegetation category is now discussed explicitly in the text. We emphasize that our main conclusion re the lake and glacier categories are unchanged by this retraining - the difference in the improvement due to the updated fields is much greater than the training noise.

	\item \textbf{Essential comment 3: MODIS observations}

	A discussion on the quality of MODIS data has been added to the text.

		\item \textbf{Essential comment 4: Errors}

	Throughout the work we use an absolute error i.e $ |{\rm LST \, predicted \, by \,  VESPER} - {\rm LST \, from \, MODIS} |$. This is now specified explicitly in the text. We have also explored the use of different error metric such as bias and RMSE, but our conclusions remain unchanged.

		\item \textbf{Essential comment 5: training and evaluating periods}

		We have trained VESPER with different input years (i.e. 2018, 2019) and results were the same. For monthly data training we agree that more data is needed and this work can be considered as our first attempt to represent and evaluate monthly lakes - the updated lake fields themselves are only for a single 12 month period. We have reworded the text accordingly to make this concession that this is our first attempt to include monthly lakes. Additionally data preparation should be more detailed - it would be useful to consider only grid cells with constant cover over the training period.

		\item \textbf{Essential comment 6: VESPER does not “beat” ERA5, it corrects ERA5}

		Fully agree, corrected in text accordingly

		\item \textbf{Essential comment 7: technical names}

		We have updated the definitions of the aggregation techniques in the text. For the $k$-nearest neighbours algorithm we feel that this is a sufficiently well known technique, common in many ML texts that it is sufficient to  specify the technique used and reference the specific Python library (RAPIDS) that we used.

			\item \textbf{Essential comment 8: predictors}

			The different VESPER models do indeed have a different vector of predictors. This has now been specified explicitly in the updated manuscript, along with better definitions with units of the various input fields (see Tables 1-3)

			\item \textbf{Essential comment 9:  vegetation and glacier updates}

			The vegetation and glacier fields were updated in proportion to the change in the lake fraction. For instance if before the fraction of the cell which is lake = 0.75 and the fraction which is glacier = 0.25, and then after the update the fraction of the cell which is lake = 0.80, the new glacier field is 0.20. This is now discussed in the updated manuscript.

			\item \textbf{Essential comment 10:  significant figures}

			Agreed. Corrected throughout manuscript

			\item \textbf{Essential comment 11: seasonal lake fraction changes and salinity.}

			This is an good point. For this work we are satisfied to consider the combined affect of monthly maps and salt lakes, since many of the locations we specifically highlight in the manuscript are saline lakes with large expected time variability in the surface water. Further work in this area is currently ongoing and we defer a more in depth, global study of saline lakes and monthly maps for the future.

			\item \textbf{Essential comment 12: What are shadows on Fig. 9-10? Please explain.}

		 Previously these shadows were confidence intervals. In the updated manuscript, with multiple VESPER trainings these shadows are the $\pm 1 \sigma$ bounds. We have specified this in the Figure captions.

	\section*{Other comments}
	All typos and editorial comments have been corrected in the updated manuscript.

\end{enumerate}

\end{document}

### 2023-05-29 docs/submission/template.tex
final documents 1

\begin{abstract}
About 2/3 of all densely populated areas (i.e. at least 300 inhabitants per km$^2$) around the globe are situated within a 9 km radius of a permanent waterbody (i.e. inland water or sea/ocean coast), since inland water sustains the vast majority of human activities. Water bodies exchange mass and energy with the atmosphere and need to be accurately simulated in numerical weather prediction and climate modelling as they strongly influence the lower boundary conditions such as skin temperatures, turbulent latent and sensible heat fluxes and moisture availability near the surface. All the non-ocean water (resolved and sub-grid lakes and coastal waters) are represented in the Integrated Forecasting System (IFS) of the European Centre for Medium-Range Weather Forecasts (ECMWF) model, by the Fresh-water Lake (FLake) parametrisation, which treats $\sim$ 1/3 of the land. It is a continuous enterprise to update the surface parametrization schemes and their input fields to better represent small-scale processes. It is, however, difficult to quickly determine both the accuracy of an updated parametrisation, and the added value gained for the purposes of numerical modelling. The aim of our work is to quickly and automatically assess the benefits of an updated lake parametrisation making use of a neural network regression model trained to simulate satellite observed surface skin temperatures. In order to rapidly assess accuracy of new surface physiography fields and if their use in the model increase predictability a neural network regression model (further in the text referred as VESPER) that can learn the mapping between a set of features x and targets y is constructed. In this case the features are the atmospheric and surface model fields (such as 2 metre temperature from ERA5 reanalysis) and the surface physiographic fields (such as orography and vegetation cover used to produce ERA5 reanalysis), see Table 1 for the full list of variables used. The target is the satellite land surface temperature (skin temperature from MODIS Aqua Day MYD11A1 v006 collection). VESPER can then make predictions about the skin temperature given a set of input variables (i.e. atmospheric and surface model fields, and surface physiographic fields). In turn, these predictions can then be compared against observations (i.e. satellite skin temperature). We deploy this tool to determine the accuracy of recent upgrades to the FLake parametrisation, namely the improved permanent lake cover and the capacity to represent seasonally varying water bodies (i.e. ephemeral lakes). We show that for grid-cells where the lake fields have been updated, the prediction accuracy in the land surface temperature improves by 0.45 K on average, whilst for the subset of points where the lakes have been exchanged for bare ground (or vice versa) the improvement is 1.12 K. We also show that updates to the glacier cover improve further the prediction accuracy by 0.14 K. The inclusion of seasonal water is shown to be particularly effective for grid points which are highly time variable, generally improving the simulation accuracy by $\sim$1 K. The neural network regression model has proven to be useful and easily adaptable to assess unforeseen impacts of ancillary datasets, also detecting inappropriate changes of high vegetation to bare ground, which would lead to decreased the skin temperature simulation accuracy by 0.49 K, proving to be a valuable support to model development.
\end{abstract}

@@

%\introduction  %% \introduction[modified heading if necessary]
\section{Introduction}
\noindent Globally, there are $\sim$ 117 million lakes - defined as inland water bodies without lateral movement of water - making up around 3.7$\%$ of the Earth's land surface \citep{Verpoorter2014}. Their distribution is highly anisotropic, with the majority of lakes located between $45-75^{\circ}$N in the Boreal and Arctic regions. Lakes are highly important from the perspective of both numerical weather prediction and climate modelling as part of the EC-Earth model. For the latter, lakes generally influence the global carbon cycle as both sinks and sources of greenhouse gases; the majority of lakes are net heterotrophic, over saturated with CO$_2$ as a result of in lake respiration and so emit carbon into the atmosphere \citep{Pace2005,Tranvik2009}.  Total CO$_2$ emission from lakes is estimated at $1.25 - 2.30$ Pg of CO2-equivalents annually \citep{DelSontro2018}, nearly $20 \%$ of global CO$_2$ fossil fuel emissions, whilst lakes account for 9-24 $\%$  of CH$_4$ emissions, the second largest natural source after wetlands \citep{Saunois2020}. These rates of greenhouse gas emission are expected to rise further if the eutrophication of the Earth's lentic systems continues. With regards to weather, freezing and melting of the lake surface modifies the radiative and conductive properties and consequently affects the heat (latent, sensible) exchange and surface energy balance \citep{Huang2019,Peng2020,Franz2018}. Considering particular examples, over Lake Victoria convective activity is suppressed during the day and peaks at night, leading to intense, hazardous thunderstorms \citep{Thiery2015,Thiery_2017}; Lake Ladoga can generate low level clouds which can cause variability in the 2m temperature of up to 10 K \citep{Eerola2014}; the Laurentian Great Lakes can cause intense winter snow storms \citep{Vavrus2013} \citep{Notaro2013}. Moreover, as a result of the increased temperatures due to climate change, lakes become more numerous due to the melting of glaciers and permafrost. Additionally, the higher temperatures mean that previously permanent lake bodies become seasonal or intermittent. There is then evidently a huge potential return in the ability to accurately model the location, morphology and properties of lakes in weather and climate models. \newline

@@

\noindent Numerical weather prediction and climate modelling are fields that are inherently linked with large datasets and complex, non-linear interactions. It is therefore an area that is particularly well placed to benefit from the deployment of machine learning algorithms. At ECMWF, advanced machine learning techniques have been used for parametrisation emulation via neural networks \citep{Chantry2021}, 4D-Var data assimilation \citep{Hatfield2021} and the post-processing of ensemble predictions \citep{Hewson2021}. Indeed, the early successes of these machine learning methods have led to the development of a 10-year roadmap for machine learning at ECMWF \citep{p19877}, with machine learning methods looking to be integrated into the operational workflow and machine learning demands considered in the procurement of HPC facilities; the ongoing development of novel computer architectures (e.g. GPU, IPU, FGPA) motivates utilizing algorithms and techniques which can efficiently take advantage of these new chips and gain significant performance returns. In this work we will demonstrate a new technique for the Verification of Earth-System ParametERisation (VESPER) based on a deep learning neural network regression model. This tool enables the accuracy of an updated water body parametrization to be rapidly and automatically assessed, and the added value that such an updated parametrization brings to be quantitatively evaluated. \newline

@@

\section{Constructing VESPER}\label{sec:2}
In order to rapidly assess accuracy of new surface physiography fields and if their use in the model increase predictability a neural network regression model (further in the text referred as VESPER) that can learn the mapping between a set of features x and targets y is constructed. In this case the features are the atmospheric and surface model fields (such as 2 metre temperature from ERA5 reanalysis) and the surface physiographic fields (such as orography and vegetation cover used to produce ERA5 reanalysis), see Table 1 for the full list of variables used. The target is the satellite land surface temperature (skin temperature from MODIS Aqua Day MYD11A1 v006 collection).VESPER can then make predictions about the skin temperature given a set of input variables (i.e. atmospheric and surface model fields, and surface physiographic fields). In turn, these predictions can then be compared against observations (i.e. satellite skin temperature) and VESPER's accuracy evaluated. By varying the number/type and values of the input features x to VESPER and observing how the accuracy of its predictions change, some conclusions on if and how features can increase predictability of an actual atmospheric model. Moreover, by isolating geographic regions where the predictions get worse with new/updated surface physiographic fields, areas where these fields might be erroneous or not informative enough can be identified. Due to the inherent stochasticity of training a neural network regression model it is also possible for different models to settle in different local minimums i.e. the network variance. To understand the significance of this, every VESPER configuration was trained four times, each time with a different random seed. \newline

@@

\subsection{Features and targets}
VESPER's input feature selection (see Table \ref{table:definitions}) followed (i) permutation importance results for atmospheric and surface model fields - only fields with the highest importance were chosen; and (ii) expert choice for surface physiographic fields - as first attempt it was decided to test current methodology for lake related information, therefore fields that could be most affected by the presence of absence of water were selected, e.g. if lake had to be removed then some other surface had to appear (like bare ground, high or low vegetation, glacier or even ocean) and surface elevation had to change. Changes to the orographic fields will have important influences on temperature through e.g. wind, solar heating etc. Lake depth changes are similarly important, influencing how a lake freezes, thaws, mixes and its overall dynamical range. VESPER's target selection followed globally available criteria and the satellite land surface temperature is quite well observed globally and with high temporal pattern (daily or even several times a day depending on the location)

@@

		\\
		Additional surface physiographic fields:               & Difference for all main surface physiographic fields between V15 and V20 field sets, \newline 
		Difference between V20 static lake cover and monthly varying lake cover (12 maps in total),\newline 
		Saline lake cover (the proportion of a grid-cell covered by saline inland water bodies; units: 0-1)
		 \\
		\hline
	\end{tabularx}
\caption{Input features used for training the NN model VESPER; atmospheric model fields (time varying) were kept the same in all simulations,  surface  physiographic  fields  (static)  were  updated  when  going  from  the  original  data  based  on GlobeCover2009/GLDBv1 (V15 field set) to GSWE/GLDBv3 (V20 field set); in brackets are variables description (where needed), short name (according to the GRIB parameter database) and units.
}
\label{table:definitions}
\end{table*}

@@

There are three main sources of data. The first is selection of surface physiographic fields from ERA5 \citep{Hersbach} and their updated versions \citep{Choulga2019,Boussetta2021,Mu2021} used as VESPER’s features.
The second is a selection of atmospheric and surface model fields from ERA5, also used as VESPER’s features. The third is day-time land surface temperature measurements from the Moderate Resolution Imaging Spectroradiometer (MODIS) onboard the Aqua satellite \citep{MODIS}, used as VESPER’s target variable.

@@

\noindent Surface physiographic fields have gridded information of the Earth’s surface properties (e.g. land-use, vegetation type and distribution) and represent surface heterogeneity in the ECLand of ECMWF’s Integrated Forecasting System (IFS). They are used to compute surface turbulent fluxes (of heat, moisture and momentum) and skin temperature over different surfaces (vegetation, bare soil, snow, interception and water) and then to calculate an area-weighted average for the grid-box to couple with the atmosphere. To trigger all different parametrization schemes the ECMWF model uses a sets of physiographic fields, that do not depend on initial condition or forecast step. Most fields are constant; surface albedo is specified for 12 months to describe the seasonal cycle. Dependent on the origin, initial data comes at different resolutions and different projections, and is then first converted to a regular latitude-longitude grid (EPSG:4326) at ~1km at Equator resolution and secondly to a required grid and resolution. Surface physiographic fields used in this work consist of orographic, land, water, vegetation, soil, albedo fields and their difference between initial V15 and updated V20 field sets (and between expanded V15X and V20X field sets). See Tables \ref{table:definitions} and \ref{tab:datasources} for the full list of surface physiographic fields and their input sources; for more details see IFS documentation \citep{IFSdocs}. As this work is focused on assessing quality of inland water information, main surface physiographic fields are lake cover (derived from land-sea mask) and lake mean depth (see Table \ref{tab:datasources}).  \newline 

@@

	\caption{List of input datasets for the surface physiographic fields for V15 and V20 field sets. V15X and V20X are identical to V15 and V20 respectively, but with the addition of saline lake cover, and monthly varying lake cover fields. \hl{TK: is glacier mask binary in V15 and 0-1 in V20?}}

@@

\noindent To generate V15 (and V15X) fractional lake cover the GlobCover2009 global map  \citep{GLOBCOVER,arino2012glcm} is used. This map has a resolution of 300m, corresponds for the year 2009 and covers latitudes 85°N-60°S; corrections outside these latitudes for the polar regions are included separately. In the Arctic no land is assumed, in the Antarctic data from the high-resolution Radarsat Antarctic Mapping Project digital elevation model version 2 (RAMP2; Liu et al., 2015) is used. To generate V20 (and V20X) fractional lake cover more resent higher resolution datasets and updated methods have been used \citep{Choulga2019}. Main data source is the Joint Research Centre (JRC) the Global Surface Water Explorer (GSWE) dataset \citep{GSWE}. GSWE is a 30m resolution dataset from Landsat 5,7 and 8, providing information on the spatial and temporal variability of surface water on the Earth since March 1984; here only permanent water type was used for lake cover generation as it was more accurate inland water distribution on the annual basis \citep{Choulga2019}. Differences between V20 and V15 lake cover fields (see Figure 1) are consistent with the latest global and regional information: (i) increase of lake fraction in V20 compared to V15 over northern latitudes is due to permafrost melt leading to a new thermokarst lake emergence, and due to higher resolution input source and its better satellite image recognition methodologies; (ii) reduction of lake fraction in V20 compared to V15 can be explained with several reasons, like anthropogenic land use change (e.g. Aral Sea, which lies across the border between Uzbekistan and Kazakhstan, has been shrinking at an accelerated rate since the 1960s and started to stabilise in 2014 with an area of 7660km2, 9 times smaller than its size in 1960. GlobCover2009 describes the Aral Sea in 1998, when it was still “only” two times smaller than its 1960 extent, whereas GSWE provides a more up to date map.), use of only permanent water (e.g. Australia, where GlobCover2009 over-represents inland water, as most of these lakes are highly ephemeral, e.g. the endorheic Kati Thanda–Lake Eyre fills only a few times per century. The GSWE updates to this region therefore include only generally permanent water, removing all seasonal and rare ephemeral water.), and change in the ocean and inland water separation algorithm (e.g. north-east of Russia). \newline

\noindent To generate V15 (and V15X) lake mean depth (see Figure \ref{example_figure_a}) the Global Lake DataBase version 1 \citep[GLDBv1;][]{Kourzeneva2012} is used. GLDBv1 has a resolution of ~1km and is based on ~13000 lakes with in situ lake depth information; outside this dataset all missing data grid-cells (i.e. over ocean and land) have 25 meter value; field aggregation to a coarser resolution is done by averaging. Overestimation of lake depth in summer season can result in strong cold biases and in winter season – lack of ice formation. To generate V20 (and V20X) lake mean depth (see Figure \ref{example_figure_b}) an updated version GLDBv3 \citep{Choulga2014} is used. GLDBv3 has the same resolution of ~1km, but is based on an increased number of lakes with in situ lake depth information by ∼1500 (in addition has bathymetry information over all Finnish navigable lakes), it introduces distinction between freshwater and saline lakes (this information is currently not used by FLake), and suggests the method to assess the depth for lakes without in situ observations using geological and climate type information; field aggregation to a coarser resolution is done by computing the most occurring value. Verification of GLDBv1 and GLDBv3 lake depths against 353 Finnish lake measurements shows that GLDBv3 exhibits a 52 \% bias reduction in mean lake depth values compared to GLDBv1 \citep{Choulga2019}. For a further details on lake distribution and depth, the representation of lakes by ECMWF in general see \cite{Choulga2019}  and \cite{Boussetta2021}. \newline

>>>>

\begin{abstract}
Water bodies exchange mass and energy with the atmosphere and need to be accurately simulated in numerical weather prediction and climate modelling as they strongly influence the lower boundary conditions such as skin temperatures, and surface fluxes of heat and moisture near the surface. All the non-ocean water (resolved and sub-grid lakes and coastal waters) are represented in the Integrated Forecasting System (IFS) of the European Centre for Medium-Range Weather Forecasts (ECMWF) model, by the Fresh-water Lake (FLake) parametrisation, which treats $\sim$ 1/3 of the land. It is a continuous enterprise to update the surface parametrization schemes and their input fields to better represent small-scale processes. It is, however, difficult to quickly determine both the accuracy of an updated parametrisation, and the added value gained for the purposes of numerical modelling. In this work a neural network regression model is developed to learn the mapping between surface data from ERA5 and MODIS satellite observations of skin temperature. Once trained, the tool can be used to verify the quality of upgrades of the land-surface parametrisation scheme since large differences between the model and the machine learning tool indicate incorrect values, which are most often caused by errors in the surface fields that are used as input to the parametrisation scheme. We deploy this model to determine the accuracy of recent upgrades to the FLake parametrisation, namely the improved permanent lake cover and the capacity to represent seasonally varying water bodies (i.e. ephemeral lakes). We show that for grid-cells where the lake fields have been updated, the prediction accuracy in the land surface temperature improves by 0.37 K on average, whilst for the subset of points where the lakes have been exchanged for bare ground (or vice versa) the improvement is 0.84 K. We also show that updates to the glacier cover improve further the prediction accuracy by 0.22 K.

\end{abstract}

@@

%\introduction  %% \introduction[modified heading if necessary]
\section{Introduction}
\noindent Globally, there are $\sim$ 117 million lakes - defined as inland water bodies without lateral movement of water - making up around 3.7$\%$ of the Earth's land surface \citep{Verpoorter2014}. Their distribution is highly non-uniform, with the majority of lakes located between $45-75^{\circ}$N in the Boreal and Arctic regions. Lakes are highly important from the perspective of both numerical weather prediction and climate modelling as part of the EC-Earth model. For the latter, lakes generally influence the global carbon cycle as both sinks and sources of greenhouse gases; the majority of lakes are net heterotrophic, over saturated with CO$_2$, as a result of in lake respiration and so emit carbon into the atmosphere \citep{Pace2005,Tranvik2009}.  Total CO$_2$ emission from lakes is estimated at $1.25 - 2.30$ Pg of CO2-equivalents annually \citep{DelSontro2018}, nearly $20 \%$ of global CO$_2$ fossil fuel emissions, whilst lakes account for 9-24 $\%$  of CH$_4$ emissions, the second largest natural source after wetlands \citep{Saunois2020}. These rates of greenhouse gas emission are expected to rise further if the eutrophication of the Earth's lentic systems continues. With regards to weather, freezing and melting of the lake surface modifies the radiative and conductive properties and consequently affects the heat (latent, sensible) exchange and surface energy balance \citep{Huang2019,Peng2020,Franz2018}. Considering particular examples, over Lake Victoria convective activity is suppressed during the day and peaks at night, leading to intense, hazardous thunderstorms \citep{Thiery2015,Thiery_2017}; Lake Ladoga can generate low level clouds which can cause variability in the 2m temperature of up to 10 K \citep{Eerola2014}; the Laurentian Great Lakes can cause intense winter snow storms \citep{Notaro2013,Vavrus2013}. Moreover, as a result of the increased temperatures due to climate change, lakes become more numerous due to the melting of glaciers and permafrost. Additionally, the higher temperatures mean that previously permanent lake bodies become seasonal or intermittent. There is then evidently a huge potential return in the ability to accurately model the location, morphology and properties of lakes in weather and climate models. \newline

@@

\noindent Numerical weather prediction and climate modelling are fields that are inherently linked with large datasets and complex, non-linear interactions. It is therefore an area that is particularly well placed to benefit from the deployment of machine learning algorithms. At ECMWF, advanced machine learning techniques have been used for parametrisation emulation via neural networks \citep{Chantry2021}, 4D-Var data assimilation \citep{Hatfield2021} and the post-processing of ensemble predictions \citep{Hewson2021}. Indeed, the early successes of these machine learning methods have led to the development of a 10-year roadmap for machine learning at ECMWF \citep{p19877}, with machine learning methods looking to be integrated into the operational workflow and machine learning demands considered in the procurement of HPC facilities. The ongoing development of novel computer architectures (e.g. GPU, IPU, FGPA) motivates utilizing algorithms and techniques which can efficiently take advantage of these new chips and gain significant performance returns. In this work we will demonstrate a new technique for the Verification of Earth-System ParametERisation (VESPER) based on a deep learning neural network regression model. This tool enables the accuracy of an updated water body parametrization to be rapidly and automatically assessed, and the added value that such an updated parametrization brings to be quantitatively evaluated. \newline

@@

\section{Constructing VESPER}\label{sec:2}
In order to rapidly assess the accuracy of new surface physiography fields and if their use in the model increase the accuracy with which we can make predictions, a neural network regression model (VESPER, hereafter) that can learn the mapping between a set of input features $x$ and targets $y$ is constructed. In this case the features are the atmospheric and surface model fields (such as 2 metre temperature from ERA5 reanalysis) and the surface physiographic fields (such as orography and vegetation cover used to produce ERA5 reanalysis). See Table \ref{table:definitions} for the full list of variables used. The target is the satellite land surface temperature (skin temperature from MODIS Aqua Day MYD11A1 v006 collection). Once trained, VESPER can then make predictions about the skin temperature given a set of input variables (i.e. atmospheric and surface model fields, and surface physiographic fields). In turn, these predictions can then be compared against observations (i.e. satellite skin temperature) and VESPER's accuracy evaluated. By varying the number, type and values of the input features to VESPER and observing how the accuracy of its predictions change, some conclusions on if and how features can increase predictability of an actual atmospheric model can be drawn. Moreover, by isolating geographic regions where the predictions get worse with new/updated surface physiographic fields, areas where these fields might be erroneous or not informative enough can be identified. Due to the inherent stochasticity of training a neural network regression model it is also possible for different models to settle in different local minimums i.e. the network variance/noise. To understand the significance of this, every VESPER configuration was trained four times, each time with a different random seed. \newline

@@

\subsection{Features and targets}
VESPER's input feature selection (see Table \ref{table:definitions}) followed (i) permutation importance results for atmospheric and surface model fields - only fields with the highest importance were chosen; and (ii) expert choice for surface physiographic fields. As a first attempt it was decided to test the current methodology for lake related information, therefore fields that could be most affected by the presence of absence of water were selected, e.g. if lake had to be removed then some other surface had to appear (like bare ground, high or low vegetation, glacier or even ocean) and surface elevation had to change. Changes to the orographic fields will have important influences on temperature through e.g. wind, solar heating etc. Lake depth changes are similarly important, influencing how a lake freezes, thaws, mixes and its overall dynamical range. VESPER's target selection followed globally available criteria and the satellite land surface temperature is quite well observed globally and with high temporal pattern (daily or even several times a day depending on the location)

@@

		\\
		Additional surface physiographic fields:               & Difference for all main surface physiographic fields between V15 and V20 field sets, \newline 
		Difference between V20 static lake cover and monthly varying lake cover (12 maps in total),\newline 
		Saline lake cover (the proportion of a grid-cell covered by saline inland water bodies; units: 0-1)
		 \\
		\hline
	\end{tabularx}
\caption{Input features used for training the neural network model VESPER; atmospheric model fields (time varying) were kept the same in all simulations,  surface  physiographic  fields  (static)  were  updated  when  going  from  the  original  data  based  on GlobeCover2009/GLDBv1 (V15 field set) to GSWE/GLDBv3 (V20 field set); in brackets are variables description (where needed), short name (according to the GRIB parameter database) and units.
}
\label{table:definitions}
\end{table*}

@@

There are three main sources of data. The first is selection of surface physiographic fields from ERA5 \citep{Hersbach} and their updated versions \citep{Choulga2019,Boussetta2021,Mu2021} used as VESPER’s features. As a shorthand we will refer to the original ERA5 physiographic fields as version ``V15" and the updated versions as ``V20". The second is a selection of atmospheric and surface model fields from ERA5, also used as VESPER’s features. The third is day-time land surface temperature measurements from the Moderate Resolution Imaging Spectroradiometer (MODIS) onboard the Aqua satellite \citep{MODIS}, used as VESPER’s target variable.

@@

\noindent Surface physiographic fields have gridded information of the Earth’s surface properties (e.g. land-use, vegetation type and distribution) and represent surface heterogeneity in the ECLand of ECMWF’s Integrated Forecasting System (IFS). They are used to compute surface turbulent fluxes (of heat, moisture and momentum) and skin temperature over different surfaces (vegetation, bare soil, snow, interception and water) and then to calculate an area-weighted average for the grid-box to couple with the atmosphere. To trigger all different parametrization schemes the ECMWF model uses a sets of physiographic fields, that do not depend on initial condition or forecast step. Most fields are constant; surface albedo is specified for 12 months to describe the seasonal cycle. Dependent on the origin, initial data comes at different resolutions and different projections, and is then first converted to a regular latitude-longitude grid (EPSG:4326) at $\sim$ 1km at equator resolution, and secondly to a required grid and resolution. Surface physiographic fields used in this work consist of orographic, land, water, vegetation, soil, albedo fields and their difference between initial V15 and updated V20 field sets. See Tables \ref{table:definitions} and \ref{tab:datasources} for the full list of surface physiographic fields and their input sources; for more details see IFS documentation \citep{IFSdocs}. As this work is focused on assessing quality of inland water information, main surface physiographic fields are lake cover (derived from land-sea mask) and lake mean depth (see Table \ref{tab:datasources}).  \newline 

@@

	\caption{List of input datasets for the surface physiographic fields for V15 and V20 field sets. V15X and V20X are identical to V15 and V20 respectively, but with the addition of saline lake cover, and monthly varying lake cover fields.}

@@

\noindent To generate V15 fractional lake cover the GlobCover2009 global map  \citep{GLOBCOVER,arino2012glcm} is used. This map has a resolution of 300m, corresponds for the year 2009 and covers latitudes 85°N-60°S; corrections outside these latitudes for the polar regions are included separately. In the Arctic no land is assumed, in the Antarctic data from the high-resolution Radarsat Antarctic Mapping Project digital elevation model version 2 (RAMP2; Liu et al., 2015) is used. To generate V20 fractional lake cover more recent higher resolution datasets and updated methods have been used \citep{Choulga2019}. The main data source is the Joint Research Centre (JRC) the Global Surface Water Explorer (GSWE) dataset \citep{GSWE}. GSWE is a 30m resolution dataset from Landsat 5,7 and 8, providing information on the spatial and temporal variability of surface water on the Earth since March 1984; here only permanent water was used for lake cover generation as it provided a more accurate inland water distribution on the annual basis \citep{Choulga2019}. Differences between V20 and V15 lake cover fields (see Figure \ref{fig:example_figure_a}) are consistent with the latest global and regional information: (i) increase of lake fraction in V20 compared to V15 over northern latitudes is due to permafrost melt leading to a new thermokarst lake emergence, and due to higher resolution input source and its better satellite image recognition methodologies; (ii) reduction of lake fraction in V20 compared to V15 can be explained with several reasons, like anthropogenic land use change (e.g. Aral Sea, which lies across the border between Uzbekistan and Kazakhstan, has been shrinking at an accelerated rate since the 1960s and started to stabilise in 2014 with an area of 7660 km$^2$, 9 times smaller than its size in 1960. GlobCover2009 describes the Aral Sea in 1998, when it was still “only” two times smaller than its 1960 extent, whereas GSWE provides a more up to date map.), use of only permanent water (e.g. Australia, where GlobCover2009 over-represents inland water, as most of these lakes are highly ephemeral, e.g. the endorheic Kati Thanda–Lake Eyre fills only a few times per century. The GSWE updates to this region therefore include only generally permanent water, removing all seasonal and rare ephemeral water.), and change in the ocean and inland water separation algorithm (e.g. north-east of Russia). \newline

\noindent To generate V15 lake mean depth (see Figure \ref{fig:example_figure_b}) the Global Lake DataBase version 1 \citep[GLDBv1;][]{Kourzeneva2012} is used. GLDBv1 has a resolution of ~1km and is based on ~13000 lakes with in situ lake depth information; outside this dataset all missing data grid-cells (i.e. over ocean and land) have 25 meter value; field aggregation to a coarser resolution is done by averaging. Overestimation of lake depth in summer season can result in strong cold biases and in winter season – lack of ice formation. To generate V20 lake mean depth an updated version GLDBv3 \citep{Choulga2014} is used. GLDBv3 has the same resolution of $\sim$1km, but is based on an increased number ($\sim$1500) of lakes with in situ lake depth information (in addition to bathymetry information over all Finnish navigable lakes), it introduces distinction between freshwater and saline lakes (this information is currently not used by FLake), and suggests the method to assess the depth for lakes without in situ observations using geological and climate type information; field aggregation to a coarser resolution is done by computing the most occurring value. Verification of GLDBv1 and GLDBv3 lake depths against 353 Finnish lake measurements shows that GLDBv3 exhibits a 52 \% bias reduction in mean lake depth values compared to GLDBv1 \citep{Choulga2019}. For a further details on lake distribution and depth, the representation of lakes by ECMWF in general see \cite{Choulga2019}  and \cite{Boussetta2021}. \newline

To expand V15 and V20 lake description (to V15X and V20X respectively, see Table \ref{tab:datasources}) their salinity and time variability information was generated. Even though static permanent water fits better to describe inland water distribution on average all year round, some areas (in Tropics especially) could  benefit from having monthly varying information as they have a very strong seasonal cycle, when size, shape and depth of a lake changes over the course of the year, leading to a significant change in modelling the lake temperature response. Similarly, saline lakes behave very differently to fresh water lakes since increased salt concentrations affect the density, specific heat capacity, thermal conductivity, and turbidity, as well as evaporation rates, ice formation and ultimately the surface temperature. These two properties of time variability and salinity are often related; it is common for saline lakes to fill and dry out over the course of the season, which naturally also affects the relative saline concentration of the lake itself. To create a monthly varying lake cover first 12 monthly fractional land-sea masks based on JRC Monthly Water History v1.3 maps for 2010-2020 were created. Since the annual lake maps were created taking into account a lot of additional sources the extra condition on the monthly maps that the monthly water is equal or greater than permanent water distribution from fractional land-sea mask is enforced. To create an inland salt lake cover map, the GLDBv3 salt lake list was used. First, in order to identify separate lakes on $\sim$ 1km resolution lake cover, small sub-grid lakes and large lake coasts are masked, i.e. grid-cells that have water fraction less than 0.25. Next, number of connected grid-cells in each lake (i.e. connected with sides only) is computed. Then only lakes that have 100 and more connected grid-cells are vectorised, as at ERA5 resolution of ∼31km the grid-cells are quite large and can include a mixture of freshwater and saline lakes. Finally, saline lake vectors are selected by filtering vectors which have no saline lake point from GLDBv3 located – in total 147 large salt lake vectors, which were further used to filter non-saline lakes at ~1km resolution lake cover, finally aggregated to ~31km resolution. In the future it is planned to revisit this field and extend the list to include additional data. Note that all non-lake related climate fields such as vegetation cover or orography were updated in V20 field set compared to V15 only in relation to the changing lake fields (i.e. if fraction of lake in the grid cell increased then other fractions like vegetation or bare ground should have increased accordingly).
